---
title: "Cjordan_DailyExercise5"
author: "Clara Jordan"
format: html
editor: visual
---

# Attach the `palmerspenguins` package

```{r}
read.csv("~/Desktop/ESS 330/Daily Assignments/palmerspenguins.csv")
```

# 1. Examine at the dataset using the ?Help page

# 2. what is the class of the penguins dataset?

```{r}
class("~/Desktop/ESS 330/Daily Assignments/palmerspenguins.csv")
```

Answer: character

# 3. what is the structure of the penguins dataset?

```{r}
structure("palmerspenguins")
```

Answer: 344 x 8

# 4. what are the dimensions of the penguins dataset?

```{r}
dim("palmerspenguins")
```

Answer: NULL

# 5. what are the column names of the penguins dataset?

Answer: species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex, year.

# 6. what type of data is `flipper_length_mm` and `Island`?

```{r}
typeof("flipper_length_mm")
typeof("Island")
```

Answer: flipper_length_mm is of type "character," Island is of type "character."

# 7. what is the mean flipper length of the penguins?

```{r}
as.logical("palmerspenguins$flipper_length_mm, na.rm =TRUE")
mean("palmerspenguins$flipper_length_mm, na.rm =TRUE")
```

Answer: I can't get past this error so I did it in Excel to get 200.915 mm.

# 8. what is the standard deviation of flipper length in the penguins?

```{r}
sd("palmerspenguins$flipper_length_mm, na.rm =TRUE")
```

Answer: I can't get past this error so I did it in Excel to get 14 mm.

# 9. what is the median body mass of the penguins?

```{r}
median("palmerspenguins$flipper_length_mm, na.rm =TRUE")
```

Answer: I don't know how to get a difference answer so I did it in Excel to get 4050 g.

# 10. what is the Island of the 100th penguin?

Answer: Dream
