---
title: "Day9/10"
author: "Clara Jordan"
format:
  html:
    self-contained: true
editor: visual
execute:
  echo: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Load Packages
```{r, eval = FALSE}
library(tidyverse)
library(tidymodels)
install.packages("visdat")
library(visdat)
library(broom)
```

### 1. Use the help (?) function to learn more about the dataset
```{r, message=FALSE, warning=FALSE}
?airquality 

  # Highlights from the dataset
    # Daily air quality measurements in New York, May to September 1973.
    # A data frame with 153 observations on 6 variables.
     # [,1]	Ozone	numeric	Ozone (ppb)
     # [,2]	Solar.R	numeric	Solar R (lang)
     # [,3]	Wind	numeric	Wind (mph)
     # [,4]	Temp	numeric	Temperature (degrees F)
     # [,5]	Month	numeric	Month (1--12)
     # [,6]	Day	numeric	Day of month (1--31)
```

### 2. Use vis_dat to check out the data. Does it need cleaning?
```{r, message=FALSE, warning=FALSE}
library(visdat)

vis_dat(airquality)

  # Answer: It looks like this dataset does need some cleaning, and has missing 
  # data for both ozone and solar radiation. All other observation categories
  # seem to have all their observations. 

  # Removal of missing data

cleaned_data <- na.omit(airquality)

vis_miss(cleaned_data) # Check to see if there is missing data still.
```

### 3. Fit a linear model to the cleaned data to predict Ozone from one of the possible predictors of your choosing. Why did you chose that variable?
```{r, message=FALSE, warning=FALSE}
model <- lm(Ozone ~ Solar.R, data = cleaned_data)

  # I chose temperature as my predictor variable because I know that 
  # tropospheric ozone tends to increase with photochemical reactions between
  # transport emissions (e.g. N2O and VOCs) and solar radiation. Thus, I think
  # there should be a direct relationship between Solar.R and Ozone. 
```

### 4. Using summary(), Does this seem like a valid model?
```{r, message=FALSE, warning=FALSE}
summary(model)

  # Answer: This does not seem like a valid model. My R^2 value is extremely low
  # at 0.1213, and the adjusted R^2 is even lower at 0.1133. The residual standard
  # error is also quite high at 31.33. Additionally, 
  # this model has very high residuals (ranging frim -48 to 119) and a large 
  # standard error of 6.7 for the intercept. 
```

### 5. Explain the R2 found in a sentence.
```{r, message=FALSE, warning=FALSE}
  # Answer: The low R^2 value, 0.1213, indicates that there is 
  # very little relationship between Solar.R and Ozone, and very little of the 
  # variance in Ozone observations is explained by solar radiation. 
```

### 6. Use broom::augment to predict the Ozone of the cleaned data.
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(visdat)
library(broom)

augmented_data <- augment(model, cleaned_data)
head(augmented_data)
```

### 7. Use ggplot to plot the actual vs predicted Ozone
```{r, message=FALSE, warning=FALSE}
correlation_value <- cor(augmented_data$Ozone, augmented_data$.fitted)
subtitle_text <- paste("Correlation:", round(correlation_value, 2))

ggplot(augmented_data, aes(x = Ozone, y = .fitted)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  labs(x = "Actual Ozone", y = "Predicted Ozone", 
       title = "Actual vs Predicted Ozone", 
       subtitle = subtitle_text)
```



