---
title: "Day11/12"
author: "Clara Jordan"
format: 
  html:
    self-contained: true
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Installing and library packages. 

library(tidyverse)
library(tidymodels)
library(visdat)
library(broom)
library(dplyr)
library(recipes)
library(lme4)
library(ggplot2)
install.packages("ggpubr", repos = "https://cran.rstudio.com/")
library(ggpubr)
```

1.  Load the airquality dataset in R. What does this dataset represent? Explore its structure using functions like `str()` and `summary()`.

```{r, eval = FALSE}
?airquality 
str(airquality)
summary(airquality)

# Highlights from the data set:
  # Daily air quality measurements in New York, May to September 1973.
  # A data frame with 153 observations on 6 variables.
    # [,1]	Ozone	numeric	Ozone (ppb)
    # [,2]	Solar.R	numeric	Solar R (lang)
    # [,3]	Wind	numeric	Wind (mph)
    # [,4]	Temp	numeric	Temperature (degrees F)
    # [,5]	Month	numeric	Month (1--12)
    # [,6]	Day	numeric	Day of month (1--31)
# This data set represents air quality measurements taken in New York.
```

2.  Perform a Shapiro-Wilk normality test on the following variables: `Ozone`, `Temp`, `Solar.R`, and `Wind.`

```{r, eval = FALSE}
airquality <- na.omit(airquality)

shapiro.test(airquality$Ozone)
shapiro.test(airquality$Temp)
shapiro.test(airquality$Solar.R)
shapiro.test(airquality$Wind)
```

3.  What is the purpose of the Shapiro-Wilk test?

```{r, eval = FALSE}
# The Shapiro-Wilk test determines whether our data are normally distributed. 
```

4.  What are the null and alternative hypotheses for this test?

```{r, eval = FALSE}
# The null hypothesis is that our data are normally distributed. 
# The alternative hypothesis is that our data are not normally distributed. 
```

5.  Interpret the p-values. Are these variables normally distributed?

```{r, eval = FALSE}
# p values indicate statistical significance if they are <0.05, and then we reject the null hypothesis. If they are >0.05 we fail to reject the null hypothesis. 

# For Ozone: p-value = 2.846e-08. This variable is normally distributed. 

# For Temperature: p-value = 0.09569. This variable is not normally distributed. 

# For Solar Radiation: p-value = 2.957e-05. This variable is normally distributed. 

# For Wind: p-value = 0.1099. This variable is not normally distributed.
```

6.  Create a new column with `case_when` tranlating the Months into four seasons (Winter (Nov, Dec, Jan), Spring (Feb, Mar, Apr), Summer (May, Jun, Jul), and Fall (Aug, Sep, Oct)).

```{r, eval = FALSE}
airquality <- airquality |>
  mutate(Season = case_when(
    Month %in% c(11, 12, 1) ~ "Winter",
    Month %in% c(2, 3, 4) ~ "Spring",
    Month %in% c(5, 6, 7) ~ "Summer",
    Month %in% c(8, 9, 10) ~ "Fall"))
```

7.  Use `table` to figure out how many observations we have from each season.

```{r, eval = FALSE}
table(airquality$Season)

# There are 52 observations from fall and 59 observations from winter. 
```

8.  Normalize the predictor variables (Temp, Solar.R, Wind, and Season) using a `recipe`

```{r, eval = FALSE}
# Creating the recipe: 
recipe <- recipe(Ozone ~ Temp + Solar.R + Wind + Season, data = airquality) %>%
  step_center(all_numeric()) %>%  
  step_scale(all_numeric())
```

9.  What is the purpose of normalizing data?

```{r, eval = FALSE}
# We normalize data so that it's on a common scale and different variables can be easily compared. Without normalizing data our results can look skewed or misrepresent reality. 
```

10. What function can be used to impute missing values with the mean?

```{r, eval = FALSE}
# step_impute_mean() can impute missing values with the mean. 
```

11. `prep` and `bake` the data to generate a processed dataset.

```{r, eval = FALSE}
prep_recipe <- prep(recipe, training = airquality)
processed_data <- bake(prep_recipe, new_data = airquality)
```

12. Why is it necessary to both `prep()` and `bake()` the recipe?

```{r, eval = FALSE}
# Prepping the recipe computes  the necessary information for preprocessing, such as means and standard deviations for scaling.
# Baking the recipe applies these computations to the data, transforming it to our needs. 
```

13. Fit a linear model using Ozone as the response variable and all other variables as predictors. Remeber that the `.` notation can we used to include all variables.

```{r, eval = FALSE}
model <- lm(Ozone ~ ., data = airquality)
```

14. Interpret the model summary output (coefficients, R-squared, p-values) in plain language

```{r, eval = FALSE}
summary(model)

# Model summary interpretation: 
  # Coefficients: The coefficients for each variable (Solar.R, Wind, Temp, etc.) indicate their relationship to Ozone. For example, with every increase of 1 unit of wind, ozone will decrease by -3.3 units. 
  # R-squared: R squared values indicate correlation between predictor and response variables in a model. The r squared for this model is 0.6. This means that there's a slight positive correlation between the predictor variables and ozone overall, but this isn't a very strong fit.
  # p-values: The p-values indicate the significance of each predictor variable. For example, the p value for day is 0.2, so this variable does not have a statistically significant effect on ozone, but temperature does because it's p value is 1.71e-10. 
```

15. Use `broom::augment` to suppliment the normalized `data.frame` with the fitted values and residuals.

```{r, eval = FALSE}
augmented_data <- augment(model, normalized_data)
```

16. Extract the residuals and visualize their distribution as a histogram and qqplot.

```{r, eval = FALSE}
residuals <- augmented_data |>
  select(.resid)

# Plot histogram and qqplot
par(mfrow = c(1, 2))  # Arrange plots side by side

# Histogram
residuals <- as.numeric(augmented_data$.resid)

hist(residuals, main = "Residuals Histogram", xlab = "Residuals", col = "pink")

# QQ plot
qqnorm(residuals)
qqline(residuals, col = "blue")
```

17. Use `ggarange` to plot this as one image and interpret what you see in them.

```{r, eval = FALSE}
hist_plot <- ggplot(data.frame(residuals), aes(x = residuals)) +
  geom_histogram(fill = "pink", bins = 20) +
  labs(title = "Residuals Histogram")

qq_plot <- ggplot(data.frame(residuals), aes(sample = residuals)) +
  geom_qq() +
  geom_qq_line() +
  labs(title = "QQ Plot of Residuals")

ggarrange(hist_plot, qq_plot, ncol = 2, nrow = 1)

# Interpretation: It looks like the histogram of the residuals is slightly right skewed, with a majority of the residuals lying around 0 and a few tailing off towards the right. This is reflected in the qq plot where most of the residuals are on the line but are tailing off as x gets larger. This means that our model is more accurate for smaller values but increasingly inaccurate as we increase the inputs. 
```

18. Create a scatter plot of actual vs. predicted values using ggpubr with the following setting:

```{r, eval = FALSE}
ggscatter(augmented_data, x = "Ozone", y = ".fitted",
          add = "reg.line", conf.int = TRUE,
          cor.coef = TRUE, cor.method = "spearman",
          ellipse = TRUE)
```

19. How strong of a model do you think this is?

```{r, eval = FALSE}
# I think this model is pretty strong. We can see that most data points cluster around the line of best fit, and that our statistical values show strong correlation (r = 0.84) and very strong statistical significance (p < 2.2e-16). 
```

20. Render your document to HTML and submit to Canvas.
