---
title: "DailyExercise5_Redo"
author: "Clara Jordan"
format:
  html:
    self-contained: true
editor: visual
execute:
  echo: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Attach the `palmerspenguins` package

```{r, message=FALSE, warning=FALSE}
library(palmerpenguins)
library(dplyr)
```

# 1. Examine at the dataset using the ?Help page

```{r, message=FALSE, warning=FALSE}
?palmerspenguins
```

# 2. what is the class of the penguins dataset?

```{r, message=FALSE, warning=FALSE}
palmerspenguins <- read.csv("palmerspenguins.csv")

class(palmerspenguins)

# Answer: The class is data.frame
```

# 3. what is the structure of the penguins dataset?

```{r, message=FALSE, warning=FALSE}
structure(palmerspenguins)

# Answer: The structure is 344 rows x 8 columns 
```

# 4. what are the dimensions of the penguins dataset?

```{r, message=FALSE, warning=FALSE}
dim(palmerspenguins)

# Answer: The dimensions are 344 rows and 8 columns
```

# 5. what are the column names of the penguins dataset?

```{r, message=FALSE, warning=FALSE}
# The column names are: species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex, year.
```

# 6. what type of data is `flipper_length_mm` and `Island`?

```{r, message=FALSE, warning=FALSE}
typeof(palmerspenguins$flipper_length_mm)
typeof(palmerspenguins$island)

# Answer: flipper_length_mm is of type "double," Island is of type "character."
```

# 7. what is the mean flipper length of the penguins?

```{r, message=FALSE, warning=FALSE}
mean(palmerspenguins$flipper_length_mm, na.rm = TRUE)
  
# Answer: The mean of flipper length is 200.92 mm
```

# 8. what is the standard deviation of flipper length in the penguins?

```{r, message=FALSE, warning=FALSE}
sd(palmerspenguins$flipper_length_mm, na.rm = TRUE)

# The standard deviation of flipper length is 14.06 mm
```

# 9. what is the median body mass of the penguins?

```{r, message=FALSE, warning=FALSE}
median(palmerspenguins$body_mass_g, na.rm = TRUE)

# Answer: The median body mass is 4050 g
```

# 10. what is the Island of the 100th penguin?

```{r, message=FALSE, warning=FALSE}
head(palmerspenguins$island, n = 100, na.rm = TRUE)

# Answer: Dream island. 
```

